---
title: Setup
---

## OS Setup

These lessons assume that you are using the freely available Visual Studio Code application with the
Benten extension along with the CWL reference runner (`cwltool`).

This tutorial requires three pieces of software to run and visualize the workflows: Docker, cwltool, and graphviz.

Please follow instructions for your OS by clicking on the relevant link below.

* [Windows Setup](#windows-setup)

* [MacOS Setup](#macos-setup)

* [Linux Setup](#linux-setup)

```{r, child = 'files/windows_setup.Rmd'}

```

Continue to [Confirm Software Installations](#confirm-software-installations) after completing the setup for Windows.

```{r, child = 'files/macos_setup.Rmd'}

```

Continue to [Confirm Software Installations](#confirm-software-installations) after completing the setup for MacOS.

```{r, child = 'files/linux_setup.Rmd'}

```

Continue to [Confirm Software Installations](#confirm-software-installations) after completing the setup for Linux.

## Confirm Software Installations

### Docker

To confirm docker is installed, run the following command to display the version number:

```bash
docker version
```

You should see something similar to the output shown below.

:::::::::::::::::::::::: spoiler

### Docker Version Output Example

```
Client: Docker Engine - Community
 Version:           20.10.13
 API version:       1.41
 Go version:        go1.16.15
 Git commit:        a224086
 Built:             Thu Mar 10 14:08:15 2022
 OS/Arch:           linux/amd64
 Context:           default
 Experimental:      true

Server: Docker Engine - Community
 Engine:
  Version:          20.10.13
  API version:      1.41 (minimum version 1.12)
  Go version:       go1.16.15
  Git commit:       906f57f
  Built:            Thu Mar 10 14:06:05 2022
  OS/Arch:          linux/amd64
  Experimental:     false
 containerd:
  Version:          1.5.10
  GitCommit:        2a1d4dbdb2a1030dc5b01e96fb110a9d9f150ecc
 runc:
  Version:          1.0.3
  GitCommit:        v1.0.3-0-gf46b6ba
 docker-init:
  Version:          0.19.0
  GitCommit:        de40ad0
```

::::::::::::::::::::::::::::::

### Confirm cwltool is installed

To confirm `cwltool` is installed, run the following command to display the version number:

```bash
cwltool --version
```

You should see something similar to the output shown below.

:::::::::::::::::::::::: spoiler

### Cwltool Version Output Example

```
/home/learner/env/bin/cwltool 3.1.20220224085855
```

::::::::::::::::::::::::

### GraphViz

To confirm `graphviz` is installed, run the following command to display the version number:

```bash
dot -V
```

You should see something similar to the output shown below.

:::::::::::::::::::::::: spoiler

### Graphviz version output example

```
dot - graphviz version 2.43.0 (0)
```

::::::::::::::::::::::::

```{r, child = 'files/containers_setup.Rmd'}

```

```{r, child = 'files/data_setup.Rmd'}

```
